A user-extensible and adaptable parser architecture

### 2008-08-19 doc/sgai-2008.tex
Corrections to the first half of the paper.

% \showgraph{filename}{caption}{label}
\newcommand{\showgraph}[3]{%
    \begin{figure}[btp]%
        \caption{#2}\label{#3}%
        \includegraphics{#1}%
    \end{figure}%
}

@@

\newcommand{\refwithlabel}[2]{%
    #1~\vref{#2}%
}
% table references, for consistent formatting.
\newcommand{\graphref}[1]{%
    \refwithlabel{fig.}{#1}%
}
\newcommand{\Graphref}[1]{%
    \refwithlabel{Figure}{#1}%
}
\newcommand{\tableref}[1]{%
    \refwithlabel{tbl.}{#1}%
}
\newcommand{\Tableref}[1]{%
    \refwithlabel{Table}{#1}%
}

@@

\abstract{%
    Some parsers need to be very precise and strict in what they parse, yet
    must allow users to easily adapt or extend the parser to parse new
    inputs, without requiring that the user have an in-depth knowledge and
    understanding of the parser's internal workings.  This paper presents a
    novel parsing architecture designed for Postfix log files which aims to
    make the process of parsing new inputs as simple as possible, enabling
    users to trivially add new rules (to parse variants of existing inputs)
    and relatively easily add new actions (to process a previously unknown
    class of input).  The architecture scales linearly as the number of
    rules and/or size of input increases, making it suitable for parsing
    large corpora or months of accumulated data.
}

@@

\newacronym{PDA}{Pushdown Automata}{}
\newacronym{FA}{Finite Automata}{}
\newacronym{CFL}{Context Free Languages}{}
\newacronym{SLCT}{Simple Logfile Clustering Tool}{}
\newacronym{SMTP}{Simple Mail Transfer Protocol}{}

@@

The architecture described herein was developed as part of a larger project
to improve anti-spam techniques by analysing the performance of the set of
filters currently in use, optimising the order and membership of the set
based on that analysis, and developing supplemental filters where there is
a deficiency.  Most anti-spam techniques are
content-based~\cite{a-plan-for-spam, word-stemming, relaxed-online-svms},
which requires accepting the mail in the first place, but rejecting mail
during the delivery attempt is preferable: senders of non-spam mail that is
mistakenly rejected will receive an immediate non-delivery notice; load is
reduced on mail servers (allowing more intensive content-based techniques
to be used); users have less mail to wade through.  Improving the
performance of anti-spam techniques which are applied when mail is being
transferred via \SMTP{}\footnotemark{} is the goal of this project, by
providing a platform for reasoning about anti-spam filters.  The approach
chosen to measure performance was to analyse the log files produced by the
mail server, rather than modifying the mail server to generate performance
statistics: this approach improves the chances of other sites testing and
utilising the software.  The need arose for a parser capable of dealing
with the great number and variety of log lines produced by
Postfix~\cite{postfix}, the \SMTP{} server in use.  The parser must be
designed so that adding support for parsing new log lines is a simple task,
as the log lines to be parsed will change over time.  The variety in log
lines occurs for several reasons:\footnotetext{Simple Mail Transfer
Protocol transfers mail from the sender to one or more recipients across
the Internet.  It is a simple, human readable, plain text protocol, making
it quite easy to test and debug problems with it.  The original protocol
definition is RFC~821~\cite{RFC821}, updated in RFC~2821~\cite{RFC2821}.}

@@

The solution developed is conceptually simple: provide a few generic
functions (\textit{actions\/}), each capable of dealing with an entire
category of log lines (e.g.\ rejecting a mail), accompanied by a multitude
of precise patterns (\textit{rules\/}), each of which matches all log lines
of a specific type, and only that type (e.g.\ rejection by a specific
DNSBL).  It is an accepted standard to separate the parsing procedure from
the declarative grammar it operates with.  Part of the novelty here is in
the way that the grammar is itself partially procedural (each action is a
separate procedure).  This architecture is ideally suited to parsing inputs
where the input is not fully understood or does not conform to a fixed
grammar: the architecture warns about unparsed inputs and other errors, but
continues on parsing as best it can, allowing the developer of a new parser
to decide which deficiencies are most important and require attention
first, rather than being forced to fix the first error which arises.

@@

    \item [Rules]  The rules are responsible for classifying or matching
        log lines, specifying the action to invoke and the data to be
        extracted from the log line.

@@

Decoupling the parsing rules from their associated actions and the
framework allows new rules to be written and tested without requiring
modifications to the parser source code, significantly lowering the barrier
to entry for casual users who need to parse new log lines, e.g.\ part-time
systems administrators attempting to reduce spam.  Decoupling the actions
from the framework simplifies both framework and actions: the framework
provides services to the actions, but does not need to perform any tasks
specific to the input being parsed; actions benefit from having services
provided by the framework, freeing them to concentrate on the task of
accurately and correctly processing the information provided by rules.

@@

There is some similarity (shown in \tableref{Similarities with ATN})
between the architecture and William Wood's Augmented Transition Networks
(ATN)~\cite{atns, nlpip}, a tool used in Computational Linguistics for
creating grammars to parse or generate sentences.  The resemblance between
ATN and the architecture is accidental, but it is interesting how two
different approaches have a similar division of responsibilities, while
having different semantics.

@@

\begin{table}[hbtp]
    \caption{Similarities with ATN}\label{Similarities with ATN}
    % The \smallskip{} below stop the text hitting the lines.
    \begin{tabular}[]{lll}
        \hline
        \noalign{\smallskip}
        ATN           & Architecture    & Similarity                          \\
        \noalign{\smallskip}
        \hline
        \noalign{\smallskip}
        Networks      & Framework       & Determines the sequence of
                                          transitions or actions which        \\
                      &                 & constitutes a valid input.          \\
        Transitions   & Actions         & Save data and impose conditions the
                                          input must meet to be               \\
                      &                 & considered valid.                   \\
        Abbreviations & Rules           & Responsible for classifying input.  \\
        \noalign{\smallskip}
        \hline
        \noalign{\smallskip}
    \end{tabular}
\end{table}

@@

Each action contains the code required to deal with a particular type of
log line, e.g.\ rejections.  The actions are parser-specific: each parser
author will need to write the actions required from scratch, unless
extending an existing parser.  In the Postfix log parser developed for this
project there are \numberOFactions{} actions and \numberOFrules{} rules,
with a very skewed distribution of rules to actions as shown in
\graphref{Distribution of rules per action}.  It is expected that parsers
utilising this architecture will have a high ratio of rules to actions,
hopefully resulting in simpler rules and clearer distinctions between
inputs.  \showgraph{build/graph-action-distribution}{Distribution of rules
per action}{Distribution of rules per action} Unsurprisingly the action
with the highest number of associated rules is \texttt{DELIVERY\_REJECTED},
the action which handles Postfix rejecting a mail; it is followed by
\texttt{SAVE\_DATA}, the action responsible for handling informative
inputs: these are log lines that supplement the data gathered from other
log lines, but that do not by themselves represent interesting events.  The
third most common action is, perhaps surprisingly, \texttt{UNINTERESTING}:
this action does nothing when executed, allowing uninteresting log lines to
be parsed without causing any effects (it does not imply that the input is
ungrammatical or unparsed).  Generally rules specifying the
\texttt{UNINTERESTING} action are rules to parse log lines which are not
associated with a specific mail, e.g.\ warnings about configuration files
changing.  The framework warns about every unparsed log line, to alert the
user that they need to alter or extend their ruleset; the Postfix log
parser successfully parses every input line in the \numberOFlogFILESall{}
log files it is currently being tested with.  The remaining actions have
only one or two associated rules: in some cases this is because there will
only ever be one type of log line for that action, e.g.\ all log lines
showing that a remote client has connected are matched by a single rule and
handled by the \texttt{CONNECT} action; in other cases the action is
required to address a deficiency in the log files, or a complication which
arises during parsing.  Taking the \texttt{CONNECT} action as an example,
it creates a new data structure for each client connecting, and saves the
data extracted by the rule (in memory, subsequently to be entered into the
database).  If the data structure already exists it is treated as a symptom
of a bug, and the action issues a warning containing the full contents of
the data structure, plus the log line which has just been parsed.

The ability to add special purpose actions to deal with difficulties and
new requirements which are discovered during parser development is one of
the strengths of this architecture: instead of writing one monolithic
function which must be modified to support new behaviour, with all the
attendant risks of adversely affecting the existing parser, when a new
requirement arises an independent action can be written to satisfy it.  In
some cases the new action will require the cooperation of other actions,
e.g.\ to set or check a flag; there is a possibility of introducing failure
by modifying existing actions, but modifications will be smaller and occur
less frequently than with a monolithic architecture, thus failures should
be less likely, and will be easier to test for and diagnose.  The parser
can be written in an object oriented style, allowing sub-classes to extend
or override actions in addition to adding new actions; because each action
is an independent function, the sub-class need only modify the action it is
interested in, rather than reproducing large chunks of functionality.

During development of the Postfix log parser it became apparent that in
addition to the obvious variety in log lines, there were a large number of
complications to be overcome.  Some were the result of deficiencies in
Postfix's logging (some of which were improved on or resolved by later
versions of Postfix); others were due to the vagaries of process
scheduling, client behaviour, and administrative actions.  All were
successfully accommodated in the Postfix log parser: adding new actions was
enough to overcome several of the complications; others required
modifications to a single existing action to work around the difficulties;
the remainder were resolved by adapting actions to pass information to
subsequent actions (via the framework), which check for the extra
information and change their behaviour appropriately.

@@

Adding new rules is simple: define all the required attributes and append
the new rule to wherever the existing rules are stored.  The Postfix log
parser supplies a utility based on \SLCT{}~\cite{slct-paper} to aid in
producing regexes from previously unparsed inputs.  Storing the rules
separately allows other approaches to rule management to be used, e.g.\
instead of manually adding new rules when unparsed log lines are seen,
machine learning techniques could be used to automatically generate new
rules, based on the existing rules and input lines.  If this approach was
taken the choice of machine learning technique would be constrained by the
size of typical data sets: the \numberOFlogFILES{} log files used to
generate statistics contain \numberOFlogLINEShuman{} log lines.  Techniques
which require the full data set when training would be impractical;
Instance Based Learning~\cite{instance-based-learning} techniques which
automatically determine which inputs from the training set are valuable and
which inputs can be discarded would probably reduce the data required to a
manageable size.  This avenue of research and development has not been
pursued by the authors, but could easily be independently undertaken.

The architecture does not try to detect overlapping rules: that
responsibility is left to the author of the rules.  Unintentionally
overlapping rules lead to inconsistent parsing and data extraction because
the first matching rule wins, and the order in which rules are tried
against each line may change between log files.  Overlapping rules are
frequently a requirement, allowing a more specific rule to match some log
lines and a more general rule to match the majority, e.g.\ separating
\SMTP{} delivery to specific sites from \SMTP{} delivery to the rest of the
world.  This can simplify both the general rule and the more specific rule;
additionally rules from different sources can be combined with a minimum of
prior cooperation or modification required.  Overlapping rules should have
a priority attribute to specify the order in which they should be used;
negative priorities may be useful for catchall rules.

Storing the rules separately to the parser allows external tools to be
written to detect overlapping rules.  Traditional regexes are equivalent in
computational power to \FA{}; it is possible to convert between regexes and
\FA{}, so regex overlap could be detected by finding a non-empty
intersection of two \FA{}\@.  One formula for \FA{} intersection is given
in~\cite{intersection-of-NFA-using-Z}: $FA1 \cap{} FA2 =
\overline{(\overline{FA1} \cup{} \overline{FA2})}$.  Perl 5.10 regexes are
more powerful than traditional regexes, e.g.\ it is possible to match
correctly balanced brackets nested to an arbitrary depth:
\newline{}\tab{}\verb!/^[^<>]*(<(?:(?>[^<>]+)|(?1))*>)[^<>]*$/!\newline{}
Perl 5.10 regexes can maintain an arbitrary state stack and are thus
equivalent in computational power to \PDA{}, so overlap detection may
require calculating the intersection of two \PDA{}\@.  \CFL{} are
equivalent in computational power to \PDA{}, and intersection of two
\CFL{s} is not closed (i.e.\ intersection can produce a language which
cannot be parsed by a \CFL{}), e.g.\ $a^{*}b^{n}c^{n} \cap a^{n}b^{n}c^{*}
\rightarrow a^{n}b^{n}c^{n}$, so \PDA{} intersection may not be meaningful
in some cases.  Detecting overlap between $n$ regexes requires
$\frac{n(n-1)}{2}$ intersections to be calculated, giving $O(n^2x)$
complexity, where $O(x)$ is the complexity of calculating intersection.
This is certainly not a task to be performed on every parsing run:
detecting overlap amongst the Postfix log parser's \numberOFrules{} rules
would require calculating \numberOFruleINTERSECTIONS{} intersections.

It is possible to define pathological regexes, which fall into two main
categories: regexes which match every input line, and regexes which consume
large amounts of CPU time during matching.  Defining a regex which matches
every input line is trivial: \verb!/^/! matches the start of every input
line.  In general a regex will not consume a large amount of CPU time when
the match is successful; usually CPU time is consumed by a regex with a lot
of alteration and variable quantifiers failing to match.  This topic is
beyond the scope of this paper, see~\cite{mastering-regular-expressions}
for in-depth discussion.

@@

    \item [Context-free rules:]  Rules are context free: they do not take
        into account past or future inputs.  In context-free
        grammar terms the parser rules could be described as:
        $\text{\textless{}input\textgreater{}} \mapsto \text{rule-1} |
        \text{rule-2} | \text{rule-3} | \dots | \text{rule-n}$

@@

    \item [Matching rules to log lines is simple:]  The first matching rule
        determines the action that will be invoked: there is no
        backtracking to try alternate rules, no attempt is made to pick a
        \textit{best\/} rule.

@@

    \item [Closer to Natural Language Processing than using a fixed
        grammar:] Unlike traditional parsers such as those used when
        compiling a programming language, this architecture does not have a
        fixed grammar specification which inputs must adhere to.  The
        architecture is capable of dealing with interleaved inputs, out of
        order inputs, and ambiguous inputs where heuristics must be applied
        --- all have arisen and been successfully dealt with in the Postfix
        log parser.

@@

@@

An important property of a parser is how parsing time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
\Graphref{Parsing time, file size, and number of lines} shows the parsing
time in seconds, file size in MB, and number of log lines in tens of
thousands, for each of the \numberOFlogFILES{} log files.  All three lines
run roughly in parallel, giving the impression that the algorithm scales
linearly with input size.  This impression is borne out by
\graphref{parsing time vs file size vs number lines factor}: the ratios are
quite tightly banded across the graph, showing that the algorithm scales
linearly.  The ratio increases (i.e.\ improves) for log files 22 and 62--68
despite their larger than average size (due to a mail forwarding loop
resulting in a greatly increased number of mails delivered and log lines
generated).

>>>>

% \showgraph{filename}{caption}{label}
\newcommand{\showgraph}[3]{%
    \begin{figure}[btp]%
        \includegraphics{#1}%
        \caption{#2}\label{#3}%
    \end{figure}%
}

@@

\newcommand{\refwithlabel}[2]{%
    #1~\vref{#2}%
}
% table references, for consistent formatting.
\newcommand{\graphref}[1]{%
    \refwithlabel{fig.}{#1}%
}
\newcommand{\Graphref}[1]{%
    \refwithlabel{Figure}{#1}%
}
\newcommand{\tableref}[1]{%
    \refwithlabel{table}{#1}%
}
\newcommand{\Tableref}[1]{%
    \refwithlabel{Table}{#1}%
}

@@

\abstract{%
    Some parsers need to be very precise and strict in what they parse, yet
    must allow users to easily adapt or extend the parser to parse new
    inputs, without requiring that the user have an in-depth knowledge and
    understanding of the parser's internal workings.  This paper presents a
    novel parsing architecture, designed for parsing Postfix log files,
    which aims to make the process of parsing new inputs as simple as
    possible, enabling users to trivially add new rules (to parse variants
    of existing inputs) and relatively easily add new actions (to process a
    previously unknown class of input).  The architecture scales linearly
    as the number of rules and/or size of input increases, making it
    suitable for parsing large corpora or months of accumulated data.
}

@@

\newacronym{PDA}{Pushdown Automata}{}
\newacronym{FA}{Finite Automata}{}
\newacronym{CFL}{Context-Free Languages}{}
\newacronym{SLCT}{Simple Logfile Clustering Tool}{}
\newacronym{SMTP}{Simple Mail Transfer Protocol}{}

@@

The architecture described herein was developed as part of a larger project
to improve anti-spam techniques by analysing the performance of the set of
filters currently in use, optimising the order and membership of the set
based on that analysis, and developing supplemental filters where there is
a deficiency.  Most anti-spam techniques are
content-based~\cite{a-plan-for-spam, word-stemming, relaxed-online-svms},
which requires accepting the mail in the first place, but rejecting mail
during the delivery attempt is preferable: senders of non-spam mail that is
mistakenly rejected will receive an immediate non-delivery notice; load is
reduced on the accepting mail server (allowing more intensive content-based
techniques to be used); users have less mail to wade through.  Improving
the performance of anti-spam techniques which are applied when mail is
being transferred via \SMTP{}\footnotemark{} is the goal of this project,
by providing a platform for reasoning about anti-spam filters.  The
approach chosen to measure performance is to analyse the log files produced
by the mail server, rather than modifying Postfix~\cite{postfix}, the
\SMTP{} server in use, to generate performance statistics: this approach
improves the chances of other sites testing and utilising the software.
The need arose for a parser capable of dealing with the great number and
variety of log lines produced by Postfix; the parser must be designed so
that adding support for parsing new inputs is a simple task, as the log
lines to be parsed will change over time.  The variety in log lines occurs
for several reasons:\footnotetext{Simple Mail Transfer Protocol transfers
mail across the Internet from the sender to one or more recipients.  It is
a simple, human readable, plain text protocol, making it quite easy to test
and debug problems with it.  The original protocol definition is
RFC~821~\cite{RFC821}, updated in RFC~2821~\cite{RFC2821}.}

@@

The solution developed is conceptually simple: provide a few generic
functions (\textit{actions\/}), each capable of dealing with an entire
category of log lines (e.g.\ rejecting a mail), accompanied by a multitude
of precise patterns (\textit{rules\/}), each of which matches all log lines
of a specific type and only that type (e.g.\ rejection by a specific
DNSBL).  It is an accepted standard to separate the parsing procedure from
the declarative grammar it operates with; part of the novelty here is in
the way that the grammar is itself partially procedural (each action is a
separate procedure).  This architecture is ideally suited to parsing inputs
where the input is not fully understood or does not conform to a fixed
grammar: the architecture warns about unparsed inputs and other errors, but
continues on parsing as best it can, allowing the developer of a new parser
to decide which deficiencies are most important and require attention
first, rather than being forced to fix the first error which arises.

@@

    \item [Rules]  The rules are responsible for classifying or matching
        log lines, specifying the action to invoke and the regex which
        matches the log line and extracts data.

@@

Decoupling the parsing rules from their associated actions and the
framework allows new rules to be written and tested without requiring
modifications to the parser source code, significantly lowering the barrier
to entry for casual users who need to parse new log lines, e.g.\ part-time
systems administrators attempting to combat and reduce spam.  Decoupling
the actions from the framework simplifies both framework and actions: the
framework provides services to the actions, but does not need to perform
any tasks specific to the input being parsed; actions benefit from having
services provided by the framework, freeing them to concentrate on the task
of accurately and correctly processing the information provided by rules.

@@

There is some similarity (shown in \tableref{Similarities with ATN})
between the architecture and William Wood's Augmented Transition Networks
(ATN)~\cite{atns, nlpip}, a tool used in Computational Linguistics for
creating grammars to parse or generate sentences.  The resemblance between
ATN and the architecture is accidental, but it is interesting how two
different approaches share a similar division of responsibilities, while
having different quite semantics.

@@

\begin{table}[hbtp]
    \caption{Similarities with ATN}\label{Similarities with ATN}
    % The \smallskip{} below stop the text hitting the lines.
    \begin{tabular}[]{lll}
        \tabletopline{}
        ATN           & Architecture    & Similarity                                \\
        \tablemiddleline{}
        Networks      & Framework       & Determines the sequence of transitions or \\
                      &                 & actions which constitutes a valid input.  \\
        Transitions   & Actions         & Save data and impose conditions the       \\
                      &                 & input must meet to be considered valid.   \\
        Abbreviations & Rules           & Responsible for classifying input.        \\
        \tablebottomline{}
    \end{tabular}
\end{table}

@@

Each action contains the code required to deal with a particular type of
log line, e.g.\ rejections.  The actions are parser-specific: each parser
author will need to write the required actions from scratch, unless
extending an existing parser.  In the Postfix log parser developed for this
project there are \numberOFactions{} actions and \numberOFrules{} rules,
with a very uneven distribution of rules to actions as shown in
\graphref{Distribution of rules per action}.  It is expected that parsers
utilising this architecture will have a high ratio of rules to actions,
hopefully resulting in simpler rules and clearer distinctions between the
inputs parsed by different rules.
\showgraph{build/graph-action-distribution}{Distribution of rules per
action}{Distribution of rules per action} Unsurprisingly the action with
the highest number of associated rules is \texttt{DELIVERY\_REJECTED}, the
action which handles Postfix rejecting a mail; it is followed by
\texttt{SAVE\_DATA}, the action responsible for handling informative
inputs: these are log lines that supplement the data gathered from other
log lines, but that do not by themselves represent interesting events.  The
third most common action is, perhaps surprisingly, \texttt{UNINTERESTING}:
this action does nothing when executed, allowing uninteresting log lines to
be parsed without causing any effects (it does not imply that the input is
ungrammatical or unparsed).  Generally rules specifying the
\texttt{UNINTERESTING} action are rules to parse log lines which are not
associated with a specific mail, e.g.\ notices about configuration files
changing.  The framework warns about every unparsed log line, to alert the
user that they need to alter or extend their ruleset; the Postfix log
parser successfully parses every input line in the \numberOFlogFILESall{}
log files it is currently being tested with.  The remaining actions only
have one or two associated rules: in some cases this is because there will
only ever be one type of log line for that action, e.g.\ all log lines
showing that a remote client has connected are matched by a single rule and
handled by the \texttt{CONNECT} action; in other cases the action is
required to address a deficiency in the log files, or a complication which
arises during parsing.  Taking the \texttt{CONNECT} action as an example,
it creates a new data structure for each client connecting, and saves the
data extracted by the rule in memory, subsequently to be entered into the
database.  If a data structure already exists for that connection it is
treated as a symptom of a bug, and the action issues a warning containing
the full contents of the existing data structure, plus the log line which
has just been parsed.

The ability to add special purpose actions to deal with difficulties and
new requirements which are discovered during parser development is one of
the strengths of this architecture: instead of writing one monolithic
function which must be modified to support new behaviour, with all the
attendant risks of adversely affecting the existing parser, when a new
requirement arises an independent action can be written to satisfy it.  In
some cases the new action will require the cooperation of other actions,
e.g.\ to set or check a flag; there is a possibility of introducing failure
by modifying existing actions, but modifications will be smaller and occur
less frequently than with a monolithic architecture, thus failures should
be less likely, and will be easier to test for and diagnose.  The parser is
written in an object oriented style, allowing sub-classes to extend or
override actions in addition to adding new actions; because each action is
an independent procedure, the sub-class need only modify the action it is
overriding, rather than reproducing large chunks of functionality.

During development of the Postfix log parser it became apparent that in
addition to the obvious variety in log lines, there were a large number of
complications to be overcome.  Some were the result of deficiencies in
Postfix's logging (some of which were improved on or resolved by later
versions of Postfix); others were due to the vagaries of process
scheduling, client behaviour, and administrative actions.  All were
successfully accommodated in the Postfix log parser: adding new actions was
enough to overcome several of the complications; others required
modifications to a single existing action to work around the difficulties;
the remainder were resolved by adapting existing actions to cooperate and
exchange extra data, changing their behaviour as appropriate.

@@

Parsing new inputs is generally achieved by creating a new rule that pairs
an existing action with a new regex.  Adding new rules is simple: define
all the required attributes and append the new rule to wherever the
existing rules are stored.  The Postfix log parser supplies a utility based
on \SLCT{}~\cite{slct-paper} to aid in producing regexes from previously
unparsed inputs.  Storing the rules separately allows other approaches to
rule management to be used, e.g.\ instead of manually adding new rules,
machine learning techniques could be used to automatically generate new
rules, based on the existing rules and input lines.  If this approach was
taken the choice of machine learning technique would be constrained by the
size of typical data sets (see \sectionref{Results}); techniques which
require the full data set when training would be impractical; Instance
Based Learning~\cite{instance-based-learning} techniques which
automatically determine which inputs from the training set are valuable and
which inputs can be discarded would probably reduce the data required to a
manageable size.  This avenue of research and development has not been
pursued by the authors, but could easily be undertaken independently.

The architecture does not try to detect overlapping rules: that
responsibility is left to the author of the rules.  Unintentionally
overlapping rules lead to inconsistent parsing and data extraction because
the first matching rule wins, and the order in which rules are tried
against each line may change between log files.  Overlapping rules are
frequently a requirement, allowing a more specific rule to match some log
lines and a more general rule to match the majority, e.g.\ separating
\SMTP{} delivery to specific sites from \SMTP{} delivery to the rest of the
world.  This can simplify both the general rule and the more specific rule;
additionally rules from different sources can be combined with a minimum of
prior cooperation or modification required.  Overlapping rules should have
a priority attribute to specify their relative ordering; negative
priorities may be useful for catchall rules.

Storing the rules separately to the parser allows external tools to be
written to detect overlapping rules.  Traditional regexes are equivalent in
computational power to \FA{}; it is possible to convert between regexes and
\FA{}, so regex overlap could be detected by finding a non-empty
intersection of two \FA{}\@.  One formula for \FA{} intersection is given
in~\cite{intersection-of-NFA-using-Z}: $FA1 \cap{} FA2 =
\overline{(\overline{FA1} \cup{} \overline{FA2})}$.  Perl 5.10 regexes are
more powerful than traditional regexes, e.g.\ it is possible to match
correctly balanced brackets nested to an arbitrary depth:
\newline{}\tab{}\verb!/^[^<>]*(<(?:(?>[^<>]+)|(?1))*>)[^<>]*$/!\newline{}
Perl 5.10 regexes can maintain an arbitrary state stack and are thus
equivalent in computational power to \PDA{}, so overlap detection may
require calculating the intersection of two \PDA{}\@.  \CFL{} are
equivalent in computational power to \PDA{}, and intersection of two
\CFL{s} is not closed, i.e.\ intersection can produce a language which
cannot be parsed by a \CFL{}, e.g.\ $a^{*}b^{n}c^{n} \cap a^{n}b^{n}c^{*}
\rightarrow a^{n}b^{n}c^{n}$, so the results of \PDA{} intersection may not
be meaningful in some cases.  Detecting overlap between $n$ regexes
requires $n(n-1)/2$ intersections to be calculated, resulting in $O(n^2x)$
complexity, where $O(x)$ is the complexity of calculating intersection.
This is certainly not a task to be performed on every parsing run:
detecting overlap amongst the Postfix log parser's \numberOFrules{} rules
would require calculating \numberOFruleINTERSECTIONS{} intersections.

It is possible to define pathological regexes, which fall into two main
categories: regexes which match every input line, and regexes which consume
excessive amounts of CPU time during matching.  Defining a regex which
matches every input line is trivial: \verb!/^/! matches the start of every
input line.  In general a regex will not consume much CPU time when the
match is successful; usually CPU time is consumed by a regex with a lot of
alteration and variable quantifiers failing to match.  This topic is beyond
the scope of this paper, see~\cite{mastering-regular-expressions} for
in-depth discussion.

@@

    \item [Context-free rules:]  Rules are context-free: they do not take
        into account past or future inputs.  In context-free grammar terms
        the parser rules could be described as:
        $\text{\textless{}input\textgreater{}} \mapsto \text{rule-1} |
        \text{rule-2} | \text{rule-3} | \dots | \text{rule-n}$

@@

    \item [Matching rules against log lines is simple:]  The first matching
        rule determines the action that will be invoked: there is no
        backtracking to try alternate rules, no attempt is made to pick a
        \textit{best\/} rule.

@@

    \item [Closer to Natural Language Processing than using a fixed
        grammar:] Unlike traditional parsers such as those used when
        compiling a programming language, this architecture does not
        require a fixed grammar specification which inputs must adhere to.
        The architecture is capable of dealing with interleaved inputs, out
        of order inputs, and ambiguous inputs where heuristics must be
        applied --- all have arisen and been successfully accommodated in
        the Postfix log parser.

@@

\label{Results}

@@

An important property of a parser is how parsing time scales relative to
input size: does it scale linearly, polynomially, or exponentially?
\Graphref{Parsing time, file size, and number of lines} shows the parsing
time in seconds, file size in MB, and number of log lines in tens of
thousands, for each of the \numberOFlogFILES{} log files.  All three lines
run roughly in parallel, giving the impression that the algorithm scales
linearly with input size.  This impression is borne out by
\graphref{parsing time vs file size vs number lines factor}: the ratios are
quite tightly banded across the graph, showing that the algorithm scales
linearly.  The ratio increases (i.e.\ improves) for log files 22 and 62--68
despite their large size (due to a mail forwarding loop
resulting in a greatly increased number of mails delivered and log lines
generated).
